Tidal streams in triaxial systems

### 2015-02-23 ms.tex
minor kvj edits

On the largest scales ($\gtrsim$30~kpc for a Milky-Way-like galaxy) in the halos of galaxies, gravitational potentials are thought to be triaxial. Despite suggestive evidence from a range of complimentary techniques, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively verified by observations. Around other galaxies, it is generally hard to measure the 3D mass profile because they are seen in projection. From the Earth's position within the Milky Way, our view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii. Unfortunately, the Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely associated with debris stripped from in-falling satellite galaxies.

@@

Tidal streams are generally modeled as 1D structures formed on regular orbits with simple internal progenitor dynamics and density evolution. Methods span a range of complexity from orbit-fitting, to \emph{Streakline} or particle-spray models, to action-space density modeling, to direct $N$-body simulations. All methods have been tested in some way on simulated observations of data and authors have mainly demonstrated their ability to recover parameters of analytic, static potential forms. Many of the methods have been shown to work in axisymmetric or otherwise simple potentials but claim to extend to more complex, generalized potentials (e.g., triaxial or multi-component potentials).

>>>>

On the largest scales in the halos of galaxies ($\gtrsim$30~kpc for a Milky-Way-like galaxy), gravitational potentials are thought to be triaxial. Despite suggestive evidence from a range of complimentary techniques, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively verified by observations. Around other galaxies, it is generally hard to measure the 3D mass profile because they are seen in projection. From the Earth's position within the Milky Way, our view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii. Unfortunately, the Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely on non-random orbits and thus contain extra information (e.g., debris stripped from in-falling satellite galaxies).

@@

Tidal streams are generally modeled as 1D structures formed on regular orbits with simple internal progenitor dynamics and density evolution. Methods span a range of complexity from orbit-fitting, to \emph{Streakline} or particle-spray models, to action-space density modeling, to N-body simulations. All methods have been tested in some way on simulated observations of data and authors have mainly demonstrated their ability to recover parameters of analytic, static potential forms. Many of the methods have been shown to work in axisymmetric or otherwise simple potentials but claim to extend to more complex, generalized potentials (e.g., triaxial or multi-component potentials).
